The \project{unRAVE} catalog

### 2016-08-23 article/unrave.tex
Some edits to the method section

\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\thecannon}{\project{The~Cannon}}
\newcommand{\rave}{\project{\acronym{RAVE}}}
\newcommand{\apogee}{\project{\acronymn{APOGEE}}}
\newcommand{\lamost}{\project{\acronymn{LAMOST}}}
\newcommand{\hipparcos}{\project{Hipparcos}}

@@

The target selection for \rave\ is based on the $I$-band apparent magnitude,
$9 < I < 12$, with a weak $J - K_s > 0.5$ cut near the disk and bulge\footnote{
However, see the works of \citet{Kunder_2016} and \citet{Kordopatis_2013} for 
subtleties of the selection function}.
The $I$ band was used because it overlaps with the wavelength range 
that \rave\ operates in:  8410--8795~\AA.  This setup includes the \ion{Ca}{2} 
near-infrared triplet.  These stellar absorption lines are dominated by pressure 
broadening, which makes them strong and therefore visible even in very low S/N data
or metal-poor stars.  This spectral region is also well-studied.  It is one of the
key setups used for the ground-based high-resolution \project{Gaia-ESO} survey
\citep{Gilmore_2012,Randich_2013}, and the \rave\ resolution and wavelength
coverage is comparable to the Radial Velocity Spectrometer on board the \project{Gaia}
space telescope \citep{Recio-Blanco_2016}.

@@

Despite the relatively low resolution and S/N of the spectra compared to other
surveys, the \rave\ data releases have demonstrated the team's enormous success 
in deriving radial velocities, stellar astrophysical parameters ($\teff$, $\logg$),
as well as individual chemical abundances \citep{Steinmetz_2006,Zwitter_2008,
Siebert_2011,Kordopatis_2013, Kunder_2016}.  In this work we make use of spectra
that has been reprocessed for the fifth \rave\ data release.  These re-processing
steps include: a detailed re-reduction of all the original data frames, with flux
variances propagated at every step; an updated continuum-normalization procedure;
as well as revised determinations of stellar radial velocities and morphological
classifications. At the end of this processing for each survey observation we are 
provided with: rest-frame wavelengths, continuum-normalized fluxes, fractional 
uncertainties in continuum-normalized flux values, as well as relevant metadata.  
We refer the reader to the official fifth data release paper of the \rave\ survey, 
as presented by \citet{Kunder_2016}, for more details of this re-processing.

@@

We chose to adopt a data-driven model --- rather than a physics-driven model ---
for this analysis.  Specifically we will use an implementation of \thecannon\
\citep{Ness_2015,Ness_2016}.  Although this choice complicated the construction 
of our model (e.g., see Section \ref{sec:the-training-set}), a data-driven model 
makes use of all available information in the spectrum, and lowers the S/N ratio 
at which systematic effects begin to dominate.  In other words, a well-constructed
data-driven model will yield more precise labels (e.g., stellar parameters and
chemical abundances) at low S/N ratios than most physics-driven models\footnote{
However, see \citet{Casey_2016a}}.  This is particularly relevant for the 
low-resolution \rave\ data analysed here, because about half of the spectra have
S/N $\lesssim 50$.

Throughout this work we adopt the assumptions and nomenclature as set out in
\citet{Casey_2016b}.  One distinction here is that the \emph{labelled} set and
the \emph{training} set in this work will be the same set.  In \citet{Casey_2016b}
the labelled set contained two subsets: the training set and a distinct
validation set, allowing us to train on the \emph{training set} and then test
on the \emph{validation set}, a subset of the labelled set.  However in this work
the labelled set is relatively small, about an order of magnitude smaller than 
what was available in \citet{Casey_2016b}, and therefore we will use all labelled 
set stars in the training step and supplement our validation experiments with 
literature comparisons.  Thus, throughout this work we will use the terms 
\emph{labelled set} and \emph{training set} interchangeably.  Below we describe 
how we constructed the training set, before we outline the model that we adopt.

@@

We sought to construct a training set consisting of stars across the main-sequence,
sub-giant, and the red giant branch.  We require stars with precisely measured
effective temperature $\teff$, surface gravity $\logg$, and elemental abundances
of Fe, Mg, Si, Ti, Ca, Al, and Ni.  This proved to be difficult because the magnitude
range of \rave\ does not overlap substantially with high-resolution spectroscopic
surveys.  The fourth internal data release of the \project{Gaia-ESO} survey includes 
giant and main-sequence stars, but only 142 overlap with \rave.  The most recent 
\project{APOGEE} data release \citep{sloan_dr13} reports labels for stars on the
giant branch and (uncalibrated values for) the main-sequence, but our tests indicated
that the main-sequence labels suffered from unexpected systematic effects.  A flat, 
then `up-turning' main-sequence is present, and the metallicity gradient trends in 
the opposite direction with respect to $\logg$ on the main-sequence (i.e., metal-poor
stars sit above the isochrones, not below).  If we move to lower-resolution studies,
there are 2,369 stars that overlap with \project{LAMOST} --- of which 2,213 have positive
S/N ratios in the $g$-band (\texttt{snrg}).  However, similar systematic effects are 
seen on the main-sequence: the labels are expectedly less precise given the lower
resolution, no detailed abundance information is available\footnote{Abundance 
information is available for \project{LAMOST} stars from \citet{Ho_2016}, but that 
sample contains only giant stars}, and there is only sparse sampling across the 
Hertsprung-Russell diagram.  The \project{GALAH} survey has a good overlap with \rave,
but results from that sample are not yet precise enough for our purposes.

These constraints forced us to construct a heterogeneous training set.  Given previous
successes in transferring high S/N ratio labels from \project{APOGEE} \citep{Ness_2015,
Ness_2016,Ho_2016,Casey_2016b}, we chose to use the 1,355 stars in the \project{APOGEE}
---\rave\ overlap sample for giant star labels in the training set.  Of these, about
900 are giants according to \project{APOGEE}.  From this sample we selected stars to
have: metallicity determinations ([Fe/H] $> -5$); high S/N ratios in both the 
\project{APOGEE} and \rave\ spectra ($>200$ in \project{APOGEE}; $>50$ in \rave); and
we further required that the \project{ASPCAP} pipeline did not report any peculiar
flags (\texttt{ASPCAPFLAG = 0}).  These restrictions left us with 357 stars along the
giant branch, with metallicities ranging from [Fe/H] $= -1.62$ to 0.32.

Assembling a training set for the main-sequence and sub-giant branch was more complex.
% training set from K2/EPIC?!!
% Identifying obviously bad stars that have been mis-classified by K2/EPIC
% Refer to the training set figure.
% Two separate models, with an overlap region.

@@

@@

@@

\bibitem[Kunder et al.(2016)]{Kunder_2016} Kunder, A., et al.\ 2016, in preparation

@@

>>>>

\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\thecannon}{\project{The~Cannon}}
\newcommand{\rave}{\project{\acronym{RAVE}}}
\newcommand{\apogee}{\project{\acronymn{APOGEE}}}
\newcommand{\lamost}{\project{\acronymn{LAMOST}}}
\newcommand{\hipparcos}{\project{Hipparcos}}
\newcommand{\epic}{\project{K2/EPIC}}

@@

The target selection for \rave\ is based on the $I$-band apparent magnitude,
$9 < I < 12$, with a weak $J - K_s > 0.5$ cut near the disk and bulge \citep{Wonjo_2016}.  
The $I$ band was used because it overlaps with the wavelength range that \rave\ 
operates in:  8410--8795~\AA.  This setup includes the \ion{Ca}{2} 
near-infrared triplet.  These stellar absorption lines are dominated by pressure 
broadening, which makes them strong and therefore visible even in very low S/N data
or metal-poor stars.  This spectral region is also well-studied.  It is one of the
key setups used for the ground-based high-resolution \project{Gaia-ESO} survey
\citep{Gilmore_2012,Randich_2013}, and the \rave\ resolution and wavelength
coverage is comparable to the Radial Velocity Spectrometer on board the \project{Gaia}
space telescope \citep{Recio-Blanco_2016}.

@@

Despite the relatively low resolution and S/N of the spectra compared to other
surveys, the \rave\ data releases have demonstrated the team's enormous success 
in deriving radial velocities, stellar astrophysical parameters ($\teff$, $\logg$),
as well as individual chemical abundances \citep{Steinmetz_2006,Zwitter_2008,
Siebert_2011,Kordopatis_2013,Kunder_2016}.  In this work we make use of spectra
that has been reprocessed for the fifth \rave\ data release.  These re-processing
steps include: a detailed re-reduction of all the original data frames, with flux
variances propagated at every step; an updated continuum-normalization procedure;
as well as revised determinations of stellar radial velocities and morphological
classifications. At the end of this processing for each survey observation we are 
provided with: rest-frame wavelengths, continuum-normalized fluxes, fractional 
uncertainties in continuum-normalized flux values, as well as relevant metadata.  
We refer the reader to the official fifth data release paper of the \rave\ survey, 
as presented by \citet{Kunder_2016}, for more details of this re-processing.

@@

We chose to adopt a data-driven model --- rather than a physics-driven model ---
for this analysis.  Specifically we will use an implementation of \thecannon\
\citep{Ness_2015,Ness_2016}.  Although this choice complicated the construction 
of our model (e.g., see Section \ref{sec:the-training-set}), a data-driven model 
makes use of all available information in the spectrum and lowers the S/N ratio 
at which systematic effects begin to dominate.  In other words, a well-constructed
data-driven model will yield more precise labels (e.g., stellar parameters and
chemical abundances) at low S/N ratios than most physics-driven models\footnote{
However, see \citet{Casey_2016a}}.  This is particularly relevant for the 
low-resolution \rave\ data analysed here, because about half of the spectra have
S/N $\lesssim 50$.

Throughout this work we adopt the assumptions and nomenclature first set out in
\citet{Ness_2015}, and extended in \citet{Casey_2016b}.  However, one distinction 
here with \citet{Caset_2016} is that in this work the \emph{labelled} set and the
\emph{training} set will be the same set.  In \citet{Casey_2016b} the labelled set
contained two subsets: the training set and a distinct validation set, allowing us
to train on the \emph{training set} and then test on the \emph{validation set}, a 
subset of the labelled set.  However in this work the labelled set is relatively 
small, about an order of magnitude smaller than what was available in 
\citet{Casey_2016b}, and therefore we will use all labelled set stars in the 
training step and supplement our validation experiments with literature comparisons.
Thus, throughout this work we will use the terms \emph{labelled set} and 
\emph{training set} interchangeably.  Below we describe how we constructed the 
training set, before we outline the model that we adopt.

@@

We sought to construct a training set consisting of stars across the main-sequence,
sub-giant, and the red giant branch.  We require stars with precisely measured
effective temperature $\teff$, surface gravity $\logg$, and elemental abundances
of Fe, Mg, Si, Ti, Ca, Al, and Ni.  This proved to be difficult because the magnitude
range of \rave\ does not overlap substantially with high-resolution spectroscopic
surveys.  The fourth internal data release of the \project{Gaia-ESO} survey includes 
giant and main-sequence stars, but only 142 overlap with \rave.  The most recent 
\project{APOGEE} data release \citep{sloan_dr13} reports labels for stars on the
giant branch and (uncalibrated values for) the main-sequence, but our tests indicated
that the main-sequence labels suffered from unexpected systematic effects.  A flat, 
then `up-turning' main-sequence is present, and the metallicity gradient trends in 
the opposite direction with respect to $\logg$ on the main-sequence (i.e., metal-poor
stars incorrectly sit above the isochrones in a classical Hertzsprung-Russell diagram)
If we consider lower-resolution studies, there are 2,369 stars that overlap with 
\project{LAMOST} --- of which 2,213 have positive S/N ratios in the $g$-band 
(\texttt{snrg}).  However, similar systematic effects are seen on the main-sequence: 
the labels are expectedly less precise given the lower resolution, no detailed 
abundance information is available\footnote{Abundance information is available for 
\project{LAMOST} stars from \citet{Ho_2016}, but that sample contains only giant stars},
and there is only sparse sampling across the Hertzsprung-Russell diagram.  The 
\project{GALAH} survey has a good overlap with \rave, but results from that sample 
are not yet precise enough for our purposes.

These constraints forced us to construct a heterogeneous training set.  Given previous
successes in transferring high S/N ratio labels from \project{APOGEE} \citep{Ness_2015,
Ness_2016,Ho_2016,Casey_2016b}, we chose to use the 1,355 stars in the \project{APOGEE}
---\rave\ overlap sample for giant star labels in the training set.  Of these, about
900 are giants according to \project{APOGEE}.  From this sample we selected stars to
have: metallicity determinations ([Fe/H] $> -5$); high S/N ratios in both the 
\project{APOGEE} and \rave\ spectra ($>200$ in \project{APOGEE}; $>50$ in \rave); and
we further required that the \project{ASPCAP} pipeline did not report any peculiar
flags (\texttt{ASPCAPFLAG = 0}).  These restrictions left us with 357 stars along the
giant branch, with metallicities ranging from $[{\rm Fe/H}] = -1.62$ to 0.32.  Our tests
showed that the metallicity range of the training set needed to extend below
$[{\rm Fe/H}] \lesssim -2$, and thus below we describe how we supplemented this
sample with giant stars from the \epic\ catalog \citep{Huber_2016}.

Assembling a training set for the main-sequence and sub-giant branch was more complex.
% training set from K2/EPIC?!!
% Identifying obviously bad stars that have been mis-classified by K2/EPIC

% Giant stars of low metallicity.

% Refer to the training set figure.
% Two separate models, with an overlap region.
% Worry later about label offsets.

@@

\subsection{The model}
\label{sec:the-model}

@@

\bibitem[Huber et al.(2016)]{Huber_2016} Huber, D., Bryson, S.~T., Haas, M.~R., et al.\ 2016, \apjs, 224, 2

@@

\bibitem[Kunder et al.(2016)]{Kunder_2016} Kunder, A., et al.\ 2016, submitted

@@

\bibitem[Wonjo et al.(2016)]{Wonjo_2016} Wonjo, J., et al.\ 2016, submitted
